A Large-Scale Study of the Use of Eval in R

### 2020-08-28 inst/paper/main.tex
AST size and VGAM
Fix introduction and background

\begin{abstract}
  The \eval function turns text into code. From the point of view of
  reasoning about software system, uses of \eval turn the program's code
  from a statically known quantity that can be reasoned about and analyzed
  into one of the inputs to the program. Thus, widespread use of \eval
  hinders our ability to provide safety and security guarantees for
  software. Understanding how \eval is used in practice is key to finding
  ways to mitigate its damage. One question that has not been studied is
  whether the usage of \eval is dependent on the application domain or other
  features of the language. This paper is a large-scale study of a corpus of
  \CorpusAllCodeRnd lines of data science libraries and end-user written in
  the R language. We performed a sophisticated dynamic analysis to confirm
  that \eval is indeed in wide spread use, and to categorized patters of
  usage. We found that \eval in this code base is more dangerous in some
  ways, and safer in others, than what was previously reported for
  JavaScript.
\end{abstract}

@@

While there has been previous work investigating the usage of \eval in web
programming and JavaScript~\cite{ecoop11}, there are no other studies that
shed light on the usage of \eval in other programming languages and
application domains.  It is thus reasonable to wonder if the results from
previous work can be carried over to other contexts. For this paper, we have
picked the data science language R~\cite{r} as our object of study. R is
both different from JavaScript in linguistic terms and in terms of its user
base. JavaScript was designed to run untrusted code in browser, while R was
design for statistical computing on a desktop. JavaScript is used as a
general purpose language by a wide community of programmers; while R is used
for scientific computing by data scientist and domain experts with, often,
limited programming experience.

@@

To understand the behavior of eval we perform a dynamic analysis and manual
code inspection. The results presented here summarize our findings.

Our infrastructure is publicly available, released in open source, and our
data, and code will be submitted to artifact evaluation.

@@

Some languages have chosen to restrict \eval. The essential difference
between languages relates to scoping of variables that can be accessed from
code executed by \eval.  Java and Julia are among the more restrictive
languages. In Java, \eval must be synthesized as a combination of dynamic
loading and reflection.\footnote{While Java does not have \eval built-in,
  one can take the input to \eval, wrap it in a static method of an
  anonymous class, generate bytecode for that class, invoke the Java class
  loader to dynamically install the new code, and then use reflection to
  call the new method.} Julia provides \eval more directly but limits its
scope to top level variables and methods. JavaScript support both a,
so-called, global \eval, and a local \eval. Where the latter has acces to
local variables of the function being executed (as well as variables of
lexically enclosing scopes). Finally languages such as R allow \eval to
manipulate variables defined in the caller environment of the function in
which \eval is invoked.

@@

Richards et al.~\cite{ecoop11} provided the first large-scale study of the
runtime behavior of \eval in JavaScript. They analyzed a corpus of the
10,000 most popular websites. They used an instrumented web browser to
gather execution traces and perform a dynamic analysis.  Their results show
that \eval is pervasive with 82\% of the most popular websites using it. The
reasons for using \eval include the desire to load code on demand (to reduce
the time it takes to display a page), the deserialization of JSON data
(about 37\% of all calls) and lightweight meta-programming to customize web
pages.  While many uses \eval were legitimate, many were unnecessary and
could be replaced with equivalent and safer code.  The reported on a
categorization of inputs to \eval that covers the vast majority of input
strings.  Restricting themselves to \eval in which all named variables refer
to the global scope, many patterns can be replaced by more disciplined
code~\cite{oopsla12b}.  The limitation of that work are that there was no
measure of code coverage, so the numbers presented can be a lower bound on
the possible behaviors.  Furthermore in 2011, JavaScript usage is quite
different from today. Their work captured very few libraries and none of the
more recent Node.js ecosystem.  More details about dynamic analysis of
JavaScript can be found in~\cite{liang}.

Wang et al.~\cite{wang} analyzed usage of dynamic features in 18 Python
programs include dynamic code generation to find if they affect file
change-proneness.  They reported that files with dynamic features are
significantly more likely to be the subject of changes, than other files.
Chen et al. looked at the correlation between code changes and dynamic
features, including \eval, in 17 Python programs~\cite{chen}. They did not
observe many uses of \eval.  Callau et al. performed an empirical study of
the usage of dynamic features in the 1,000 largest Smalltalk projects in the
Squeaksource repository, accounting for more than 4 million lines of
code~\cite{oscar}. While \eval itself is not present, Smalltalk has a rich
reflective interface that gives programmer great flexibility. Their work was
based on a static analysis.

@@

To roughly estimate the complexity of the code being evaluated,
Fig.~\ref{fig:ast-size} shows the distribution of the abstract syntax tree
size of the \c{expr} argument to \eval. The majority of inputs across core
and packages (\NbAstOnePercent of \eval calls and \NbAstOneCallSitePercent
of call sites) have size 1. That is to say, the code being evaluated
consists of a single variable name or constant. A function call needs at
least three nodes and a list such as \c{c(1, 2, 3)} counts for 4
nodes. Inputs smaller than \AstSizeNineFive nodes account for 95\% of all
calls. Sizes in core are larger than in packages. Most large AST originate
from library \c{VGAM}, which is used for statistical modelling. It uses
\eval to evaluate large matrices with values in a given environment.

>>>>

\begin{abstract}
  The \eval function turns text into code. From the point of view of
  reasoning about software system, uses of \eval turn the program's code
  from a statically known quantity that can be reasoned about and analyzed
  into one of the inputs to the program. Thus, widespread use of \eval
  hinders our ability to provide safety and security guarantees for
  software. Understanding how \eval is used in practice is key to finding
  ways to mitigate its damage. One question that has not been studied is
  whether the usage of \eval is dependent on the application domain or other
  features of the language. This paper is a large-scale study of a corpus of
  \CorpusAllCodeRnd lines of data science libraries and end-user code written in
  the R language. We performed a sophisticated dynamic analysis to confirm
  that \eval is indeed in wide spread use, and categorized patterns of
  usage. We found that \eval in this code base is more dangerous in some
  ways, and safer in others, than what was previously reported for
  JavaScript.
\end{abstract}

@@

While there has been previous work investigating the usage of \eval in web
programming and JavaScript~\cite{ecoop11}, there are no other studies that
shed light on the usage of \eval in other programming languages and
application domains.  It is thus reasonable to wonder if the results from
previous work can be carried over to other contexts. For this paper, we have
picked the data science language R~\cite{r} as our object of study. R is
both different from JavaScript in linguistic terms and in terms of its user
base. JavaScript was designed to run untrusted code in browser, while R was
designed for statistical computing on a desktop. JavaScript is used as a
general purpose language by a wide community of programmers; while R is used
for scientific computing by data scientists and domain experts with, often,
limited programming experience.

@@

To understand the behavior of \eval we perform a dynamic analysis and manual
code inspection. The results presented here summarize our findings.

Our infrastructure is publicly available, released in open source, and our
code, and data will be submitted for artifact evaluation.

@@

Some languages have chosen to restrict \eval. The essential difference
between languages relates to scoping of variables that can be accessed from
code executed by \eval.  Java and Julia are among the more restrictive
languages. In Java, \eval must be synthesized as a combination of dynamic
loading and reflection.\footnote{While Java does not have \eval built-in,
  one can take the input to \eval, wrap it in a static method of an
  anonymous class, generate bytecode for that class, invoke the Java class
  loader to dynamically install the new code, and then use reflection to
  call the new method.} Julia provides \eval more directly but limits its
scope to top level variables and methods. JavaScript supports both a,
so-called, global \eval, and a local \eval, where the latter has acces to
local variables of the function being executed (as well as variables of
lexically enclosing scopes). Finally, languages such as R allow \eval to
even manipulate variables defined in the caller environment of the function in
which \eval is invoked.

@@

Richards et al.~\cite{ecoop11} provided the first large-scale study of the
runtime behavior of \eval in JavaScript. They analyzed a corpus of the
10,000 most popular websites. They used an instrumented web browser to
gather execution traces and performed a dynamic analysis.  Their results show
that \eval is pervasive with 82\% of the most popular websites using it. The
reasons for using \eval include the desire to load code on demand (to reduce
the time it takes to display a page), the deserialization of JSON data
(about 37\% of all calls) and lightweight meta-programming to customize web
pages.  While many uses \eval were legitimate, many were unnecessary and
could be replaced with equivalent and safer code.  They reported on a
categorization of inputs to \eval that covers the vast majority of input
strings.  Restricting themselves to \eval in which all named variables refer
to the global scope, many patterns could be replaced by more disciplined
code~\cite{oopsla12b}.  The limitation of that work are that there was no
measure of code coverage, so the numbers presented can be a lower bound on
the possible behaviors.  Furthermore in 2011, JavaScript usage was quite
different from today. Their work captured very few libraries and none of the
more recent Node.js ecosystem.  More details about dynamic analysis of
JavaScript can be found in~\cite{liang}.

Wang et al.~\cite{wang} analyzed usage of dynamic features in 18 Python
programs include dynamic code generation to find if they affect file
change-proneness.  They reported that files with dynamic features are
significantly more likely to be the subject of changes, than other files.
Chen et al. looked at the correlation between code changes and dynamic
features, including \eval, in 17 Python programs~\cite{chen}. They did not
observe many uses of \eval.  Callau et al.~\cite{oscar} performed an empirical study of
the usage of dynamic features in the 1,000 largest Smalltalk projects in the
Squeaksource repository, accounting for more than 4 million lines of
code~\cite{oscar}. While \eval itself is not present, Smalltalk has a rich
reflective interface that gives programmer great flexibility. Their work was
based on a static analysis.

@@

To roughly estimate the complexity of the code being evaluated,
Fig.~\ref{fig:ast-size} shows the distribution of the abstract syntax tree
size of the \c{expr} argument to \eval. The majority of inputs across core
and packages (\NbAstOnePercent of \eval calls and \NbAstOneCallSitePercent
of call sites) have size 1. That is to say, the code being evaluated
consists of a single variable name or constant. A function call needs at
least three nodes and a list such as \c{c(1, 2, 3)} counts for 4
nodes. Inputs smaller than \AstSizeNineFive nodes account for 95\% of all
calls. Sizes in core are larger than in packages. Most large AST originate
from library \c{VGAM}, which is used for statistical modeling, and also calls functions using \eval from core. It uses
\eval to evaluate large matrices with values in a given environment.
